---
title: Mermaid Diagrams in RMarkdown
author: rodney
date: '2021-03-30'
slug: mermaid-diagrams-in-rmarkdown
categories:
  - Programming
  - Visualization
  - R
tags:
  - tips
subtitle: ''
summary: ''
authors: []
lastmod: '2021-03-30T06:32:46-04:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


A nice tool to have at your disposal is a method for quickly making a connected graph diagram in your markdown.  It may be a workflow you are trying to show or something an org chart or whatever.  One thing that I've recently learned to through [Obsidian](https:://obsidian.md) as I develop my second brain (I call it my *brainforest*) is that you can easily insert mermaid charts into markdown and have them render properly. In R, it is the same, though we have a few tricks we need to use so that it looks nice.


```{r}
library(DiagrammeR)
```


A `mermaid` chart is a simple textual representation of nodes and connections denoted as a string.  There is a top down graph-like flowchart.

```{r}
mermaid("graph TD; 
	A[Start] --> B{Is it?}; 
	B -- Yes --> C[OK]; 
	C --> D[Rethink]; 
	D --> B; 
	B -- No --> E[End];")
```

The thing that I find a bit of a pain (in RMarkdown though apparently not here in blogdown) is that the box around the plot in RMarkdown is a bit too large.  To correct for that, we need to add to the function call that the diagram needs to take up 100% of the space.

```{r}
mermaid("graph LR; 
A(Process 1)-->B(Process 2);
B-->C(Analysis);
C-->D(Visualization);
D-->E(QA/QC);",
        width="100%",height="100%")
```


So that it all fits together.
