---
title: "Unneceessary Output by dplyr"
description: |
  I love tidyverse but cannot understand how the authors disregard meaningless output so poorly.  Here is how to quiet some of that extra cruft when you load in tidyverse
author:
  - name: Rodney Dyer
    url: https://dyerlab.org
    affiliation: Center for Environmental Studies
    affiliation_url: https://ces.vcu.edu
    orcid_id: 0000-0003-4707-3453
date: 2022-02-01
preview: featured.png
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
library( tidyverse )
options( tidyverse.quiet=FALSE)

```



I love to use tidyverse when working on data but there is a bad habit that the authors have by spamming your output with messages they thing are informative.  We need to have some basic ways to turn off this output if we are to use things link `distil` to publish actual manuscripts directly instead of just simple markdown documents.

For example, consider the case when you load in the library `tidyverse`.  You get all this cruft which 99% of the time is just annoyring.

```{r eval=FALSE}
library( tidyverse )
```

<pre>
Registered S3 methods overwritten by 'dbplyr':
   method         from
   print.tbl_lazy     
   print.tbl_sql      
 ── Attaching packages ───────────────────────────────────────────────────────────── tidyverse 1.3.1 ──
 ✓ ggplot2 3.3.5     ✓ purrr   0.3.4
 ✓ tibble  3.1.6     ✓ dplyr   1.0.7
 ✓ tidyr   1.2.0     ✓ stringr 1.4.0
 ✓ readr   2.1.2     ✓ forcats 0.5.1
 ── Conflicts ──────────────────────────────────────────────────────────────── tidyverse_conflicts() ──
 x dplyr::filter() masks stats::filter()
 x dplyr::lag()    masks stats::lag()
</pre>

That is now what we want!  There are several other options (like asking about conflicts, warning that `group_by` is not carried over after a summarize, etc), and we get it already.  To turn these off put the following into your `.Rprofile` file at the root of your home directory and it will make a beginning stab at it!

```bash
tidyverse.quiet = TRUE
dplyr.summarise.inform = FALSE
```