---
title: "Craft + RStudio + Distill = Post "
description: |
  This is the R side of the Craft -> Website through RStudio and Distill.  I've set up a workflow here to take posts exported in markdown from Craft to be imported into this site's workflow.
author:
  - name: Rodney Dyer
    url: https://dyerlab.org
date: 2021-12-21
preview: emily-reimer-W3RjW1rnHN0-unsplash.jpg
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

So the first thing to do was to set up a staging place to save raw markdown files.  Once I had that, I moved the rendering of this website from the "push the button on the build pane in RStudio to "run this script."  

This script is in the rood directory of the site and does the following steps:

1. finds all markdown files in the `_toImport` folder.  For each file it:
  - Loads in the file
  - Uses the first line as the metadata `title:`
  - Use the first paragraph as the metadata `description:` 
  - Creates a new `distill::create_post()` with title and modification date for the markdown file.
  - Removes boilerplate stuff form new post and puts in content from the markdown file.
  - Adds content to the metadata (extended author info and featured image components).
  - Renames post to `index.Rmd` to make it easier to link in (I wish this was an option in distill).
  - Renders the markdown.
  - Deletes original markdown file.
2. After finishing up, it then renders the entire site.

Here is the script.

```{r}
#' This file is how I'll build the site.
#' 
rm( list=ls() )


# load in and convert any of the markdown stuff in _toImport into 
#  distill articles.
for( file in list.files("_toImport", pattern = "*.md", full.names = TRUE) ) { 
  
  # find previous stuff
  raw <- readLines( file )
  date <- stringr::str_split( file.info(file)$mtime , 
                              " ", 
                              simplify = TRUE)[,1] 
  title <- stringr::str_remove( raw[1], "# ") 
  description <- raw[3]
  contents <- paste( raw[ -1 ], collapse="\n")
  
  # make new post  will create 
  filetitle <- gsub("[[:punct:][:blank:]]+", "-", title ) 
  filetitle <- tolower(filetitle)
  orig_title <- paste( "_posts/", 
                       date, "-", 
                       filetitle, "/", 
                       filetitle,".Rmd", 
                       sep="" )
  
  distill::create_post( filetitle, 
                        date = date,
                        edit = FALSE )
  
  raw_post <- readLines( orig_title )
  raw_post[2] <- paste( "title: \"", title, "\"" )
  raw_post[4] <- paste( "  ", description, sep=" ")
  raw_post[7] <- "    url: https://dyerlab.org\n    affiliation: Center for Environmental Studies\n    affiliation_url: https://ces.vcu.edu\n    orcid_id: 0000-0003-4707-3453"
  raw_post[8] <- paste( raw_post[8], "preview: featured.png", sep="\n")
  raw_post <- c( raw_post[1:13], contents ) 
  raw_post_output <- paste( raw_post, collapse="\n")
  
  new_file <- paste( "_posts/", 
                     date, "-", 
                     filetitle, "/", 
                     "index.Rmd", sep="" )
  writeLines( raw_post_output, 
              con = new_file )
  
  # clean up the posts
  unlink( orig_title )
  unlink( file ) 
  
  rmarkdown::render( new_file )
}

# clean up any of the crap on the site and render the whole thing.
system( "find . -type f -iname '.DS_Store' -delete")
rmarkdown::render_site(".")
```


What I do not do is push through all the git actions to upload it.  I think I'll just do that manually.

[Here](../2021-12-21-craft-app-rstudio-distill-blogging/) is an example of the first Craft $\to$ Distill posting.

